### 2017-07-23 src/5_3_Proposal.tex
algoritmos revisados falta actualizar graficos

\noindent where $\mathbf{S}{t}$ and $\mathbf{W}{t}$ are define as:

@@

The algorithm~\ref{alg:proposal} shows OVECM using three different
methods for geting model parameters: OLS, RR, AAR which leads to three different
algorithms OVECM-OLS, OVECM-RR, OVECM-AAR:

\begin{algorithm}[ht]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{y}$: matrix with $N$ input vectors and $l$ time series\\
$p$: number of past values \\
method: \{'OLS','RR','AAR'\} \\
$\lambda$: regularization parameter (only requires for RR and AAR methods) \\
$L$: window size of data($L<N$) \\
$\text{mean\_error}$: MAPE threshold \\
$n$: windows size to obtain in-sample MAPE \\
\ENSURE  $\,$ \\
$\{\mathbf{y}_{\text{pred}}[L+1],\dots, \mathbf{y}_{\text{pred}}[N]\}$: model predictions 
\STATE solver = new \texttt{Solver}(method,$\lambda$) \\
\FOR { $t =1$ to $N-L$ }
    \STATE $\mathbf{Y} \gets \mathbf{y}[t:t+L-1]$
    \STATE $\mathbf{y}_t \gets \mathbf{y}[t+L]$
	\IF {$t = 1$}
	    \STATE{$v \gets \texttt{getJohansen}(\mathbf{Y},p)$}
	    \STATE{$[\mathbf{A} \quad \mathbf{B}] \gets
        \texttt{vecMatrix}(\mathbf{Y},p,v)$}
        \STATE \texttt{solver.init\_model($\mathbf{A},\mathbf{B}$)} 
    \ENDIF
	\STATE{[$\mathbf{A} \quad \mathbf{B} \quad \mathbf{a}_t \quad \mathbf{a}_L \quad \mathbf{b}_t \quad
    \mathbf{b}_L ] \gets
    \texttt{vecMatrixOnline}(\mathbf{Y},\mathbf{y}_t,p,v)$}
    \STATE $[\mathbf{X} \quad \mathbf{y}_{\text{pred}}[t]] \gets \texttt{solver.regression}
    (\mathbf{A},\mathbf{B},\mathbf{a}_t,\mathbf{b}_t)$
    \STATE $\mathbf{Y}_{\text{pred}} = \mathbf{AX}[-n:]+\mathbf{Y}[-n-1:-1]$
    \STATE $\mathbf{e} = \texttt{mape}(\mathbf{Y}_{\text{pred}},\mathbf{Y}[-n:])$
    \IF {$\texttt{mean}(\mathbf{e}) > \text{mean\_error}$}
	    \STATE{$v \gets \texttt{getJohansen}(\mathbf{Y},p)$}
	    \STATE{$\mathbf{A} \gets
        \texttt{vecMatrixUpdate}(\mathbf{A},\mathbf{Y},p,v)$}
        \STATE \texttt{solver.init\_model($\mathbf{A},\mathbf{B}$)} 
	    \STATE{$[\mathbf{a}_t \quad \mathbf{b}_t] \gets
        \texttt{vecMatrixOnline}(\mathbf{Y},\mathbf{y}_t,p,v)$}
        \STATE $[\mathbf{X} \quad \mathbf{y}_{\text{pred}}[t]] \gets \texttt{solver.regression}
        (\mathbf{A},\mathbf{B},\mathbf{a}_t,\mathbf{b}_t)$
    \ENDIF
\STATE{$[\mathbf{A} \quad \mathbf{B}] \gets
\texttt{matrixUpdate}(\mathbf{A},\mathbf{B},\mathbf{a}_t,\mathbf{b}t)$}
\ENDFOR
\end{algorithmic}
\caption{OVECM: Online VECM}
\label{alg:proposal}
\end{algorithm}

The proposal considers the following:

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by the
Johansen method considering the trace statistic test at 95\% level of
significance. This procedure is called at the first step of the algorithm and
every time in-sample MAPE ($\mathbf{e}$) exceeds threshold error defined (mean\_error).
\item The function \texttt{vecMatrix} returns VECM
matrices $\mathbf{A}(t),\mathbf{B}(t)$ shown in equation~(\ref{eq:notation}). This
method is only required at the first step.
\item The function \texttt{vecMatrixOnline} returns new rows $\mathbf{a}_t^\top$ and
$\mathbf{b}_t^\top$ given new input data $\mathbf{y}_t$.
\item The Solver class (see algorithm \ref{alg:solver}) obtain VECM parameters  
using RR, AAR or OLS methods.
\item The function \texttt{vecMatrixUpdate} updates matrix $\mathbf{A}$ when new
cointegration vectors are required (matrix $\mathbf{B}$ is not affected). Model
parameters $\mathbf{X}$ is also updated later.
%\item The number of cointegration vectors is set as a parameter (aun no lo
%coloco en el algoritmo).
%\item In order to set VECM parameter: $L$ and $p$, we use the Akaike Information
%Criterion (AIC). RR parameter $\lambda$ was done by cross-validation.
%\item Obtain VECM parameter $\mathbf{X}(t)$ in equation \ref{eq:optsolSLAAR} requires
%to calculate the inverse of matrix $\mathbf{S}(t+1)$ which is an expensive
%routine.However, since we already know $\mathbf{S}(t)^{-1}$ we can use
%Sherman-Morrison-Woodbury twice~\ref{eq:SMW}.
\end{itemize}

\begin{algorithm}[ht]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
method: \{'OLS','RR','AAR'\} \\
$\mathbf{A}$: VECM design matrix \\
$\mathbf{B}$: VECM dependant variables \\
$\mathbf{a}_t$: new row of $\mathbf{A}$ \\
$\mathbf{b}_t$: new row of $\mathbf{B}$ \\
$\gamma$: regularization parameter \\
$n$: windows size to obtain in-sample MAPE \\
\ENSURE  $\,$ \\
$\mathbf{X}$: regression solution \\
$\mathbf{e}$: in-sample MAPE \\
%\quad \\
%\texttt{init}($L,\lambda$)
%\STATE self.L = L
%\STATE self.lambda = $\lambda$
\quad \\
\texttt{init\_model}($\mathbf{A},\mathbf{B}$)
\STATE $ [m \quad n] = \text{size}(\mathbf{A}) $ 
\STATE $\mathbf{S} = \displaystyle \sum_{i=1}^m \mathbf{a}_i \mathbf{a}_i^\top + \gamma \mathbb{I}$
\STATE $\mathbf{W} = \displaystyle \sum_{i=1}^m \mathbf{a}_i \mathbf{b}_i$
\quad \\
\texttt{regression}($\mathbf{A},\mathbf{B},\mathbf{a_t},\mathbf{b_t}$) \\
\IF {method == 'OLS'}
        \STATE $\mathbf{X}=(\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top
        \mathbf{B}$
\ELSIF {method == 'RR'}
        \STATE $\mathbf{X} = \mathbf{S}^{-1} \mathbf{W} $
        \STATE $\mathbf{S} = \mathbf{S}+
        \mathbf{a}_t \mathbf{a}_t^\top-
        \mathbf{a}_1 \mathbf{a}_1^\top$
        \STATE $\mathbf{W} = \mathbf{W} + \mathbf{a}_t \mathbf{b}_t$
\ELSIF {method == 'ARR'}
        \STATE $\mathbf{S} = \mathbf{S}+
        \mathbf{a}_t \mathbf{a}_t^\top-
        \mathbf{a}_1 \mathbf{a}_1^\top$
        \STATE $\mathbf{X} = \mathbf{S}^{-1} \mathbf{W} $
        \STATE $\mathbf{W} = \mathbf{W} + \mathbf{a}_t \mathbf{b}_t$
\ENDIF
\STATE $\mathbf{y}_{\text{pred}} = \mathbf{X}^\top \mathbf{a}_t$
\end{algorithmic}
\caption{Solver Class for Regression Methods}
\label{alg:solver}
\end{algorithm}

@@

The proposal was compared against an optimal online algorithm which is a time varying
VECM (OOVECM). Algorithm ~\ref{alg:OOVECM} shows this modification:

\begin{algorithm}[ht]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{y}$: matrix with $N$ input vectors and $l$ time series\\
$p$: number of past values \\
$L$: sliding window size ($L<N$) \\
\ENSURE  $\,$ \\
$\{\Delta \mathbf{y}_{\text{pred}}[L+1],\dots,\Delta \mathbf{y}_{\text{pred}}[N]\}$: model predictions 
\FOR { $i =0$ to $N-L$ }
    \STATE $\mathbf{y}_i \gets \mathbf{y}[i:i+L]$
	\STATE{$v \gets \texttt{getJohansen}(\mathbf{y}_i,p)$}
	\STATE{$[\mathbf{A} \quad \mathbf{B}] \gets
    \texttt{vecMatrix}(\mathbf{y}_i,p,v)$}
    \STATE $\mathbf{X} \gets \text{OLS} (\mathbf{A},\mathbf{B})$%\mathbf{(A^\top A)^{-1}A^\top B}$
    \STATE $\Delta \mathbf{Y}_{\text{true}}[i] \gets \mathbf{B}[-1,:]$
    \STATE $\Delta \mathbf{Y}_{\text{pred}}[i] \gets \mathbf{A}[-1,:] \times \mathbf{X}$
\ENDFOR
    \STATE $\text{MAPE} \gets \texttt{mape}(\Delta \mathbf{Y}_{\text{true}}, \Delta
    \mathbf{Y}_{\text{pred}})$
\end{algorithmic}
\caption{OOVECM: Optimal Online Vector Error Correction Model}
\label{alg:OOVECM}
\end{algorithm}

>>>>

\noindent where $\mathbf{S}{t}$ and $\mathbf{W}{t}$ are defined as:

@@

The algorithm~\ref{alg:proposal} shows OVECM using three different
methods for geting model parameters: OLS, RR, AAR which leads to three different
algorithms OVECM-OLS, OVECM-RR, OVECM-AAR.
The proposal considers the following:

\begin{itemize}
\item The function \texttt{getJohansen} returns cointegration vectors given by the
Johansen method considering the trace statistic test at 95\% level of
significance. This procedure is called at the first step of the algorithm and
every time in-sample MAPE ($\mathbf{e}$) exceeds threshold error defined (mean\_error).
\item The function \texttt{vecMatrix} returns VECM
matrices $\mathbf{A}(t),\mathbf{B}(t)$ shown in equation~(\ref{eq:notation}). This
method is only required at the first step.
\item The function \texttt{vecMatrixOnline} returns new rows $\mathbf{a}_t^\top$ and
$\mathbf{b}_t^\top$ given new input data $\mathbf{y}_t$.
\item The Solver class (see algorithm \ref{alg:solver}) obtain VECM parameters  
using RR, AAR or OLS methods.
\item The function \texttt{vecMatrixUpdate} updates matrix $\mathbf{A}$ when new
cointegration vectors are required (matrix $\mathbf{B}$ is not affected). Model
parameters matrix $\mathbf{X}$ is also updated later.
\item The number of cointegration vectors is set as a parameter % agregar a alg
\item RR parameter $\lambda$ was done by cross-validation. %agregar cap timeseries
\end{itemize}

\newpage
\begin{algorithm}[ht]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{y}$: matrix with $N$ input vectors and $l$ time series\\
$p$: number of past values \\
method: \{'OLS','RR','AAR'\} \\
$\lambda$: regularization parameter (only requires for RR and AAR methods) \\
$L$: window size of data($L<N$) \\
$\text{mean\_error}$: MAPE threshold \\
$n$: windows size to obtain in-sample MAPE \\
\ENSURE  $\,$ \\
$\{\mathbf{y}_{\text{pred}}[L+1],\dots, \mathbf{y}_{\text{pred}}[N]\}$: model predictions 
\STATE solver = new \texttt{Solver}(method,$\lambda$) \\
\FOR { $t =1$ to $N-L$ }
    \STATE $\mathbf{Y} \gets \mathbf{y}[t:t+L-1]$
    \STATE $\mathbf{y}_t \gets \mathbf{y}[t+L]$
	\IF {$t = 1$}
	    \STATE{$v \gets \texttt{getJohansen}(\mathbf{Y},p)$}
	    \STATE{$[\mathbf{A} \quad \mathbf{B}] \gets
        \texttt{vecMatrix}(\mathbf{Y},p,v)$}
        \STATE \texttt{solver.init\_model($\mathbf{A},\mathbf{B}$)} 
    \ENDIF
	\STATE{[$\mathbf{A} \quad \mathbf{B} \quad \mathbf{a}_t \quad \mathbf{a}_L \quad \mathbf{b}_t \quad
    \mathbf{b}_L ] \gets
    \texttt{vecMatrixOnline}(\mathbf{Y},\mathbf{y}_t,p,v)$}
    \STATE $[\mathbf{X} \quad \mathbf{y}_{\text{pred}}[t]] \gets \texttt{solver.regression}
    (\mathbf{A},\mathbf{B},\mathbf{a}_t,\mathbf{b}_t)$
    \STATE $\mathbf{Y}_{\text{pred}} = \mathbf{AX}[-n:]+\mathbf{Y}[-n-1:-1]$
    \STATE $\mathbf{e} = \texttt{mape}(\mathbf{Y}_{\text{pred}},\mathbf{Y}[-n:])$
    \IF {$\texttt{mean}(\mathbf{e}) > \text{mean\_error}$}
	    \STATE{$v \gets \texttt{getJohansen}(\mathbf{Y},p)$}
	    \STATE{$\mathbf{A} \gets
        \texttt{vecMatrixUpdate}(\mathbf{A},\mathbf{Y},p,v)$}
        \STATE \texttt{solver.init\_model($\mathbf{A},\mathbf{B}$)} 
	    \STATE{$[\mathbf{a}_t \quad \mathbf{b}_t] \gets
        \texttt{vecMatrixOnline}(\mathbf{Y},\mathbf{y}_t,p,v)$}
        \STATE $[\mathbf{X} \quad \mathbf{y}_{\text{pred}}[t]] \gets \texttt{solver.regression}
        (\mathbf{A},\mathbf{B},\mathbf{a}_t,\mathbf{b}_t)$
    \ENDIF
\STATE{$[\mathbf{A} \quad \mathbf{B}] \gets
\texttt{matrixUpdate}(\mathbf{A},\mathbf{B},\mathbf{a}_t,\mathbf{b}t)$}
\ENDFOR
\end{algorithmic}
\caption{OVECM: Online VECM}
\label{alg:proposal}
\end{algorithm}

\newpage
\begin{algorithm}[ht!]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
method: \{'OLS','RR','AAR'\} \\
$\mathbf{A}$: VECM design matrix \\
$\mathbf{B}$: VECM dependant variables \\
$\mathbf{a}_t$: new row of $\mathbf{A}$ \\
$\mathbf{b}_t$: new row of $\mathbf{B}$ \\
$\gamma$: regularization parameter \\
$n$: windows size to obtain in-sample MAPE \\
\ENSURE  $\,$ \\
$\mathbf{X}$: regression solution \\
$\mathbf{e}$: in-sample MAPE \\
%\quad \\
%\texttt{init}($L,\lambda$)
%\STATE self.L = L
%\STATE self.lambda = $\lambda$
\quad \\
\texttt{init\_model}($\mathbf{A},\mathbf{B}$)
\STATE $ [m \quad n] = \text{size}(\mathbf{A}) $ 
\STATE $\mathbf{S} = \displaystyle \sum_{i=1}^m \mathbf{a}_i \mathbf{a}_i^\top + \gamma \mathbb{I}$
\STATE $\mathbf{W} = \displaystyle \sum_{i=1}^m \mathbf{a}_i \mathbf{b}_i$
\quad \\
\texttt{regression}($\mathbf{A},\mathbf{B},\mathbf{a_t},\mathbf{b_t}$) \\
\IF {method == 'OLS'}
        \STATE $\mathbf{X}=(\mathbf{A}^\top \mathbf{A})^{-1}\mathbf{A}^\top
        \mathbf{B}$
\ELSIF {method == 'RR'}
        \STATE $\mathbf{X} = \mathbf{S}^{-1} \mathbf{W} $
        \STATE $\mathbf{S} = \mathbf{S}+
        \mathbf{a}_t \mathbf{a}_t^\top-
        \mathbf{a}_1 \mathbf{a}_1^\top$
        \STATE $\mathbf{W} = \mathbf{W} + \mathbf{a}_t \mathbf{b}_t$
\ELSIF {method == 'ARR'}
        \STATE $\mathbf{S} = \mathbf{S}+
        \mathbf{a}_t \mathbf{a}_t^\top-
        \mathbf{a}_1 \mathbf{a}_1^\top$
        \STATE $\mathbf{X} = \mathbf{S}^{-1} \mathbf{W} $
        \STATE $\mathbf{W} = \mathbf{W} + \mathbf{a}_t \mathbf{b}_t$
\ENDIF
\STATE $\mathbf{y}_{\text{pred}} = \mathbf{X}^\top \mathbf{a}_t$
\end{algorithmic}
\caption{Solver Class for Regression Methods}
\label{alg:solver}
\end{algorithm}

@@

\newpage
The proposal was compared against an optimal online algorithm which is a time varying
VECM (OVECM-OPT). Algorithm ~\ref{alg:OOVECM} shows this modification:

\begin{algorithm}[ht!]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{y}$: matrix with $N$ input vectors and $l$ time series\\
$p$: number of past values \\
$L$: sliding window size ($L<N$) \\
\ENSURE  $\,$ \\
$\{\Delta \mathbf{y}_{\text{pred}}[L+1],\dots,\Delta \mathbf{y}_{\text{pred}}[N]\}$: model predictions 
\FOR { $i =0$ to $N-L$ }
    \STATE $\mathbf{y}_i \gets \mathbf{y}[i:i+L]$
	\STATE{$v \gets \texttt{getJohansen}(\mathbf{y}_i,p)$}
	\STATE{$[\mathbf{A} \quad \mathbf{B}] \gets
    \texttt{vecMatrix}(\mathbf{y}_i,p,v)$}
    \STATE $\mathbf{X} \gets \text{OLS} (\mathbf{A},\mathbf{B})$%\mathbf{(A^\top A)^{-1}A^\top B}$
    \STATE $\Delta \mathbf{Y}_{\text{true}}[i] \gets \mathbf{B}[-1,:]$
    \STATE $\Delta \mathbf{Y}_{\text{pred}}[i] \gets \mathbf{A}[-1,:] \times \mathbf{X}$
\ENDFOR
    \STATE $\text{MAPE} \gets \texttt{mape}(\Delta \mathbf{Y}_{\text{true}}, \Delta
    \mathbf{Y}_{\text{pred}})$
\end{algorithmic}
\caption{OVECM-OPT: Optimal Online Vector Error Correction Model}
\label{alg:OOVECM}
\end{algorithm}
